---
title: 'Analysis: Epidemiologic investigation into FMD in Maasai Mara Kenya'
author: "Grewar, J,D."
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(epiR)
library(dplyr)
library(terra)
library(lme4)


```

```{r import_datasets, include=TRUE, echo = FALSE}
df <- data.frame(readxl::read_excel("D:/OneDrive/jDATA/git/saehp-shiny-server/jdata/epicpd/botswanaVS/www/coursematerial/tabular/nthiwa_kenya_seroprev.xlsx"))
baseURL = "D:/OneDrive/jDATA/git/saehp-shiny-server/jdata/epicpd/botswanaVS/www/coursematerial"
#Call generic functions
source(paste0(baseURL,"/functions/01_functions.R"))

prev.overall = epi.prev(pos = sum(df$fmd_exp_st), tested = nrow(df), 
         se = SeParrTwoTest(0.864, 0.864), 
         sp = SpParrTwoTest(0.981, 0.974), method = "wilson",
         units = 100, conf.level = 0.95)
```

## Results

A total of `r nrow(df)` from `r length(unique(df$hse_id))` animals were were tested during the survey. Male animals made up `r 100- round(sum(df$anim_sex)/nrow(df)*100,2)`% of the population with the remaining `r round(sum(df$anim_sex)/nrow(df)*100,2)`% female. An average of `r mean(df %>% group_by(hse_id) %>% tally() %>% .$n)` animals were tested per herd.

### Prevalence

On overall apparent prevalence of `r paste(round((prev.overall[[1]][1]),2))`% (95% CI `r paste(round((prev.overall[[1]][2]),2))`% - `r paste(round((prev.overall[[1]][3]),2))`%) of FMD was found throughout the study. When accounting for the sensitivity (86.4% and 86.4% respectively) and specificity (98.1% and 97.4% respectively) of the tested used, with an interpretation in parallel, the true prevalence was estimated at `r paste(round((prev.overall[[2]][1]),2))`% (95% CI `r paste(round((prev.overall[[2]][2]),2))`% - `r paste(round((prev.overall[[2]][3]),2))`%).

#### Zonal Differences in Prevalence

To evaluate differences between zones 20 km buffers surrounding Maasai Mara reserve were established. Figure 1 below depicts the study site - Narok region of Kenya. The 20 km zones are depcited as well as the animals sampled and classified by their FMD outcome during the survey. 

```{r map_zones, include=TRUE, echo=FALSE}
colnames(df)[10] <- "lat"
colnames(df)[11] <- "lon"
samplLoc = terra::vect(df,  crs="+proj=longlat +datum=WGS84")
samplLoc$fmd_exp_stFCT = factor(samplLoc$fmd_exp_st, levels = c(0,1), labels = c("Negative","Positive"))

#Use local files to connect - importShapefilesOnline functions would also work

narok = importShapefilesLocal(filePlusExt = "narok.zip")
masaiMara = importShapefilesLocal(filePlusExt = "masaiMara.zip")

# Buffer section ####
buffers = c(20000, 40000, 60000) 
buffersManuscript = lapply(buffers,
                           function(x) 
                             crop(narok, 
                                  buffer(masaiMara, x)
                                  )
                           )

b20 = buffersManuscript[[1]]
b40 = buffersManuscript[[2]]
b60 = buffersManuscript[[3]]

b20$class = "20km"
b40$class = "40km"
b60$class = "60km"

b60 = erase(b60, b40)
b40 = erase(b40, b20)

bAll = rbind(b20, b40, b60)
bAll = bAll[,"class"] #retain just the class tag
# Buffers end



#cattleDensAfrica = terra::rast(paste0(baseURL,"/gis/rasters/densCattleAfrica.tif"))
plot(narok, pax=list(sides=1:2))
plot(samplLoc, "fmd_exp_stFCT", col=c("green", "red"), add = TRUE, legend = TRUE,
     plg=list(x=36.2, y=-1.75, cex=1.2))
plot(masaiMara, "AREANAME", density=10, angle=45, col="orange", add = TRUE, legend = TRUE,
     plg=list(x=36.2, y=-1.5, cex = 1.2))
plot(bAll, "class", alpha = 0.4, col = c("pink","purple","red"), add = TRUE, legend = TRUE,
     plg=list(x=34.5, y=-1.75, cex=1.2))
```
```{r Table_Prev, echo=FALSE}
df$study_str <- factor(df$study_str)
levels(df$study_str) <- c("Zone 3", "Zone 2","Zone 1")
outputList = list()

for (i in unique(df$study_str)){
  datatemp = df[df$study_str == i,]
  
  outputList[[i]] = epi.prev(pos = sum(datatemp$fmd_exp_st), tested = nrow(datatemp), 
         se = SeParrTwoTest(0.864, 0.864), 
         sp = SpParrTwoTest(0.981, 0.974), 
         method = "wilson",
         units = 100, 
         conf.level = 0.95)
}

tablePrev = plyr::ldply(outputList, data.frame) #nice way to fit a list to dataframe
tablePrev = tablePrev %>% mutate(ap = paste0(round(tablePrev$ap.est, 2),"% (",
                                             round(tablePrev$ap.lower, 2),"% - ",
                                             round(tablePrev$ap.upper, 2),"%)"),
                                 tp = paste0(round(tablePrev$tp.est, 2),"% (",
                                             round(tablePrev$tp.lower, 2),"% - ",
                                             round(tablePrev$tp.upper, 2),"%)")
                                 ) %>% 
  select(.id, ap, tp)

colnames(tablePrev) <- c("Zone","Apparant Prevalence","True Prevalence")
  
knitr::kable(tablePrev, caption = "Zone based prevalence accounting for test uncertainty - 95% CI's shown")

```
#### Risk factor evaluation

Univariate risk factor analysis has been performed accounting for individual herd as a random effect. Significant factors were later included in a multivariate analysis. 

##### Sex of animal

```{r univariate, echo = FALSE, include=TRUE}
df$anim_sex = factor(df$anim_sex, levels = c(0,1), labels = c("male","female"))

glmer.sex = glmer(
  fmd_exp_st~anim_sex+(1|hse_id),
  data=df,
  family = binomial) #the response variable is a success or a failure - binomial is appropriate option

# summary(glmer.sex)

glmer.sex.se <- sqrt(diag(vcov(glmer.sex))) #compare with the summary standard errors

# table of estimates with 95% CI
glmer.sex.tab <- cbind(Est = fixef(glmer.sex), LL = fixef(glmer.sex) - 1.96 * glmer.sex.se, UL = fixef(glmer.sex) + 1.96 *
                glmer.sex.se)

outcome = as.data.frame(exp(glmer.sex.tab))

knitr::kable(outcome, caption = "Univariate analysis - Sex of animal - reported by Odd's ratio with upper and lower CI")

```


